A Neural Network Alternative to Convolutive Audio Models for Source Separation

### 2017-06-12 paper/paper.tex

% Template for ICASSP-2010 paper; to be used with:
%          mlspconf.sty  - ICASSP/ICIP LaTeX style file adapted for MLSP, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{amsmath,graphicx,mlspconf}
\usepackage{amsfonts}

@@

% Title.
% ------
\title{A Neural Network Alternative to Convolutive Audio Models for Source Separation}
%
% Single address.
% ---------------
\name{Author(s) Name(s) omitted for double blind review\thanks{Thanks to XYZ agency for funding.}}
\address{Author Affiliation(s) omitted for double blind review}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%   Email A-B
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D
%   Email C-D}
%
\begin{document}
%\ninept
%

@@

@@

As discussed above, the encoder part acts as the inverse filter to the decoder. We can think of a CNN as a filter with a finite impulse response. As known from DSP theory, the inverse of an FIR filter is an IIR filter. To the best of our knowledge, this results in a novel RNN-CNN auto-encoder architecture. In the remainder of the paper, we will refer to this auto-encoder as a Recurrent-Convolutive Auto-encoder~(RCAE). The encoder RNN equation is obtained as follows:

\begin{align}
  \mathbf Z(k_1, t, i) =& g\Bigg( \sum_{k_2}\mathbf W^i(k_1,k_2 ) \mathbf Z(k_2, t-1, i) + \notag \\
  & \hspace{2cm} \sum_l \mathbf U^i(k_1,l) \mathbf X(l, t) \Bigg), \\
  \mathbf H(i, t) =& \sum_{k_1} \mathbf Z^i(k_1, t)
\end{align}
where

>>>>

% Template for ICASSP-2010 paper; to be used with:
%          mlspconf.sty  - ICASSP/ICIP LaTeX style file adapted for MLSP, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{amsmath,graphicx,mlspconf}
\usepackage{amsfonts}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,calc}

@@

% Title.
% ------
\title{A Neural Network Alternative to Convolutive Audio Models for Source Separation}
%
% Single address.
% ---------------
\name{Author(s) Name(s) omitted for double blind review\thanks{Thanks to XYZ agency for funding.}}
\address{Author Affiliation(s) omitted for double blind review}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%   Email A-B
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D
%   Email C-D}
%

\tikzset{
block/.style = {draw, fill=white, rectangle, minimum height=3em, minimum width=3em},
tmp/.style  = {coordinate}, 
sum/.style= {draw, fill=white, circle, node distance=1cm},
input/.style = {coordinate},
output/.style= {coordinate},
pinstyle/.style = {pin edge={to-,thin,black}
}
}

\begin{document}
%\ninept
%

@@

\begin{figure*}
  \centering
  \begin{tikzpicture}[auto, node distance=2cm,>=latex']

      \node [,name=input] (input) {$X$};
      \coordinate [, name=rinput, right of=input, node distance=2cm ](rinput) {};

      \node [, right of=rinput] (controller) {$\vdots$};
      \node [block, above of=controller,node distance=1.3cm] (up){$\mathbf W^\ddagger_1$};
      \node [block, below of=controller,node distance=1.3cm] (rate) {$\mathbf W^\ddagger_K$};
      \node [sum, right of=controller,node distance=2cm] (sum2) {CT};

      \node [block, right of=sum2,node distance=2cm] (system) {$\frac{a_{\beta 2}}{s+a_{\beta 1}}$};
      \node [output, right of=system, node distance=2cm] (output) {};
      \node [tmp, below of=controller] (tmp1){$H(s)$};

      \draw [->] (rinput) -- (controller); 
      \draw [->] (controller) -- (sum2);
      \draw [->] (sum2) -- node{$U(s)$} (system);
     % \draw [->] (system) -- node [name=y] {$Y(s)$}(output);
      \draw [->] (rinput) |- (rate);
      \draw [->] (rate) -| (sum2);
      \draw [->] (rinput) |- (up);
      \draw [->] (up) -| (sum2);
      % \draw [->] (y) |- (tmp1)-| node[pos=0.99] {$-$} (sum1);
      %\draw [->] (extra)--(sum2);
      %\draw [->] ($(0,1.5cm)+(extra)$)node[above]{$d_{\beta 2}$} -- (extra);
    \end{tikzpicture}
  \caption{Block Diagram of CNN-CNN Autoencoder} 
  \label{CNN-CNN Auto Encoder}
\end{figure*}

@@

%As discussed above, the encoder part acts as the inverse filter to the decoder. We can think of a CNN as a filter with a finite impulse response. As known from DSP theory, the inverse of an FIR filter is an IIR filter. To the best of our knowledge, this results in a novel RNN-CNN auto-encoder architecture. In the remainder of the paper, we will refer to this auto-encoder as a Recurrent-Convolutive Auto-encoder~(RCAE).

In this section, the goal is to construct a recurrent encoder analogous to the convolutive encoder we discussed in the previous section. The potential gain of using a recurrent encoder over a finite length convolutive encoder is due to the fact that a recurrent filter in theory can capture arbitrarily long temporal dependencies. From a signal processing point of view, the motivation for going with a recurrent filter is that, the inverse of an finite length filter (the convolutive basis in the encoder) is given by a recurrent filter. We explore the idea, and obtain source separation results on the same dataset.

The way we go about building an encoder is by passing the input through $K$ separate recurrent neural networks (RNNs) (Note that $K$ was the number of filters/components in the convolutive model). The $k$'th RNN recursion in the encoder is given by the following equation:

\begin{align}
  \mathbf Z(k_1, t, k) =& \tanh \Bigg( \sum_{k_2 = 1}^{K_{in}} \mathbf W^k(k_1,k_2 ) \mathbf Z(k_2, t-1, k) + \notag \\
  & \hspace{0cm} \sum_l \mathbf U^k(k_1,l) \mathbf X(l, t) \Bigg),\; k \in \{1,\dots,K\}, 
\end{align}
where $\mathbf Z(:, t, k) \in \mathbb R^{K_{in}}$ denotes the latent state vector of the $k$'th RNN at time $t$. We denote the hidden state dimensionality of each RNN with $K_{in}$. The recurrent and projection matrices of the $k'th$ RNN are respectively denoted with $\mathbf W^k$, and $\mathbf U^k$. Note that although the given recursion corresponds to the vanilla-RNN architecture, there is no restriction on the RNN architecture choice. In our experiments, we have used the LSTM architecture [cite]. After going through the RNN recursions, the encoder output $\mathbf H(i, t)$ is obtained by summing the RNN outputs over the first dimension:
\begin{align}
  \mathbf H(i, t) = \sum_{k_1=1}^{K_{in}} \mathbf Z(k_1, t, i)
\end{align}
